### 2024-04-07 code/analysis-code/ML-analysis.qmd
part 4

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

## Introduction

This script uses tuberculosis(tb) as the outcome of interest and fits the following models to the analysis data:

* Null
* LASSO
* Random Forest
* Support Vector Machine

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Each model will follow this process:
1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

```{r seed=123}
defaultW <- getOption("warn") 
options(warn = -1) 
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(yardstick)
library(probably) #install.packages('probably')
tidymodels_prefer()
library(kernlab) 
options(warn = defaultW)

```

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","data2.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)
head(madaproject)
colnames(madaproject)
```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
# evaluate class imbalance 
madaproject %>% count(tb)
```
I see that i have more people without tb

Let me look at the levels 
```{r}
levels(madaproject$tb)
```
I see that NO is predicted first but I want yes

```{r}
# If "YES" is not already the first level, use relevel() to interchange the levels
madaproject$tb <- relevel(madaproject$tb, ref = "YES")

# Check the levels again to confirm the change
levels(madaproject$tb)
```
Now I see that YES is the first level

Now I will split the data into training and testing 
```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)

# Data splits , train, test
data_split <- initial_split(madaproject, prop = 0.7)
train <- training(data_split)
test <- testing(data_split)
ntrain = nrow(train)
ntest = nrow(test)
```

FITTING OF ALL MODELS 
Fitting the null model
```{r}
#Fit a null model to the outcome tb
logistic_model <- logistic_reg() %>% set_engine("glm")
null_model <- logistic_model %>% fit(tb ~ 1, data = train)

#Compute acurrancy  for the null model
metrics_null <- null_model %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, .pred_class)
#Print accurancy for the null model
print(metrics_null)

```
LASSO MODEL TUNING

```{r}
set.seed(rngseed)
# Make sure you set reference level (the outcome you are NOT interested in)
levels(madaproject$tb)

data_cv10 <- vfold_cv(train, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(tb ~ ., data = train) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_recipe(logistic_rec) %>%
    add_model(logistic_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
lasso_grid <- grid_regular(
  penalty(range = c(-3, 1)), 
  levels = 30)

tune_outputlasso <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE),
  grid = lasso_grid # lasso penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_outputlasso) + theme_classic()

# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_outputlasso, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the lowest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_lasso <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
fit(data = train)

```

```{r}
set.seed(rngseed)
# Predict probabilities
lassomodel_probs <- final_fit_lasso %>%
  predict(new_data = train, type = "prob")  
lassomodel_with_probs <- bind_cols(train, lassomodel_probs)

#Calculate ROC curve
roc_curve_lasso <- roc_curve(data = lassomodel_with_probs, truth = tb, .pred_YES)

Lasso_ROCCURVE <- autoplot(roc_curve_lasso)
Lasso_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "Lasso_ROCCURVE.png")
ggsave(filename = figure_file, plot=Lasso_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(final_fit_lasso, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matlasso <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
lassoconf_table<- as.data.frame(conf_matlasso$table)

# Create a flextable from the data frame
#lassoconf_table <- flextable(conf_dflasso)

# Print the table
print(lassoconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "lassoconf_table.rds")
saveRDS(lassoconf_table, file = summarytable_file)
```

Now I will do Random Forest
```{r}
# Random Forest Model Spec
rf_spec_tune <- rand_forest() %>%
    set_mode("classification") %>%
    set_engine("ranger") # or any other suitable engine

# Recipe
rf_rec <- recipe(tb ~ ., data = train)

# Workflow (Recipe + Model)
rf_wf <- workflow() %>% 
    add_recipe(rf_rec) %>%
    add_model(rf_spec_tune)

# Tune Model (trying a variety of values of mtry and trees)
rf_grid <- grid_regular(
  mtry(range = c(1,7)), # adjusting based on number of predictors
  trees(range = c(1, 21)), # adjust the range as needed
  levels = 5 # adjust the number of levels as needed
)

tune_outputrf <- tune_grid( 
  rf_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc),
  control = control_resamples(save_pred = TRUE),
  grid = rf_grid, # parameter grid defined above
)
print(tune_outputrf)
# Visualize Model Evaluation Metrics from Tuning
#autoplot(tune_outputrf)

# Fit Final Model
final_fit_rf <- finalize_workflow(rf_wf, select_best(tune_outputrf, "roc_auc")) %>% 
    fit(data = train)

```

```{r}

set.seed(rngseed)
# Predict probabilities
rfmodel_probs <- final_fit_rf %>%
  predict(new_data = train, type = "prob")  
rfmodel_with_probs <- bind_cols(train, rfmodel_probs)

#Calculate ROC curve
roc_curve_rf <- roc_curve(data = rfmodel_with_probs, truth = tb, .pred_YES)

rf_ROCCURVE <- autoplot(roc_curve_rf)
rf_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "rf_ROCCURVE.png")
ggsave(filename = figure_file, plot=rf_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(final_fit_rf, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matrf <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
rfconf_table <- as.data.frame(conf_matrf$table)

# Create a flextable from the data frame
#rfconf_table <- flextable(conf_dfrf)

# Print the table
print(rfconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "rfconf_table.rds")
saveRDS(rfconf_table, file = summarytable_file)
```

Using Support Vector Machine 
```{r}
# Support Vector Machine Model Spec
svm_spec_tune <- svm_poly() %>%
    set_mode("classification") %>%
    set_engine("kernlab") # or any other suitable engine for SVM

# Recipe
svm_rec <- recipe(tb ~ ., data = train)

# Workflow (Recipe + Model)
svm_wf <- workflow() %>% 
    add_recipe(svm_rec) %>%
    add_model(svm_spec_tune)

# Tune Model (trying a variety of values of C and degree for polynomial kernel)
svm_grid <- grid_regular(
  cost(range = c(0.1, 10)), # adjust the range as needed
  degree(range = c(1, 5)), # adjust the range as needed
  levels = 5 # adjust the number of levels as needed
)

tune_outputsvm <- tune_grid( 
  svm_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc),
  control = control_resamples(save_pred = TRUE),
  grid = svm_grid # parameter grid defined above
)

# Fit Final Model
svmfinal_fit <- finalize_workflow(svm_wf, select_best(tune_outputsvm , "roc_auc")) %>% 
    fit(data = train)

```

```{r}
set.seed(rngseed)
# Predict probabilities
svmmodel_probs <- svmfinal_fit %>%
  predict(new_data = train, type = "prob")  
svmmodel_with_probs <- bind_cols(train, svmmodel_probs)

#Calculate ROC curve
roc_curve_svm <- roc_curve(data = rfmodel_with_probs, truth = tb, .pred_YES)

svm_ROCCURVE <- autoplot(roc_curve_svm )
svm_ROCCURVE
# Save the table as a file
figure_file = here("results","figures", "svm_ROCCURVE.png")
ggsave(filename = figure_file, plot=svm_ROCCURVE) 
```

```{r}
set.seed(rngseed)
# Compute confusion matrix
predictions <- predict(svmfinal_fit, new_data = train)
pred_train <- bind_cols(train, predictions)
conf_matsvm <- pred_train %>%
  conf_mat(truth = tb, estimate = .pred_class)

# Convert confusion matrix to data frame
svmconf_table <- as.data.frame(conf_matsvm$table)

# Create a flextable from the data frame
#svmconf_table <- flextable(conf_dfsvm)

# Print the table
print(svmconf_table)

# Save the table as a file
summarytable_file <- here("results","tables", "svmconf_table.rds")
saveRDS(svmconf_table, file = summarytable_file)

```
Organising results of accuracy and RMSE

```{r}
# Null model
# Compute the accuracy and AUC for null model
null_acc <- null_model %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

null_auc <-  null_model %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#LASSO model
# Compute the accuracy and AUC for Lasso
lasso_acc <- final_fit_lasso %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

lasso_auc <-  final_fit_lasso %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
rf_acc <- final_fit_rf %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

rf_auc <-  final_fit_rf %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Support Vector Machine
# Compute the accuracy and AUC for SVM
svm_acc <- svmfinal_fit %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

svm_auc <-  svmfinal_fit %>%
  predict(train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)

```

```{r}
# Print the results
print(null_acc)
print(lasso_acc)
print(rf_acc)
print(svm_acc)
print(null_auc)
print(lasso_auc)
print(rf_auc)
print(svm_auc)
```
```{r}
# Create a data frame with model names, accuracy, and AUC
model_perfomance_train <- data.frame(
  Model = c("Null Model", "Lasso Model", "Random Forest Model", "Support Vector model"),  # Replace with your model names
  Accuracy = c(0.5781931	, 0.7831776	, 0.8255452	, 0.7838006	),  # Replace with your accuracy values
  AUC = c(0.5	, 0.8707883	, 0.8973205		, 0.8618063		)  # Replace with your AUC values
)

# Print the results
print(model_perfomance_train)

# Save the table as a file

summarytable_file <- here("results","tables", "model_perfomance_train.rds")
saveRDS(model_perfomance_train, file = summarytable_file)

```

MODEL EVALUTIONS
I will fit the models again using test data and look at the results

```{r}

# Null model
# Compute the accuracy and AUC for null model
null_acc <- null_model %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

null_auc <-  null_model %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#LASSO model
# Compute the accuracy and AUC for model 2
lasso_acc <- final_fit_lasso %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

lasso_auc <-  final_fit_lasso %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
rf_acc <- final_fit_rf %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

rf_auc <-  final_fit_rf %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
#Random model
# Compute the accuracy and AUC for model 2
svm_acc <- svmfinal_fit %>% 
  predict(test) %>% 
  bind_cols(test) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))

svm_auc <-  svmfinal_fit %>%
  predict(test, type = "prob") %>%
  bind_cols(test) %>%
  roc_auc(truth = tb, .pred_YES)

```

```{r}
# Print the results
print(null_acc)
print(lasso_acc)
print(rf_acc)
print(svm_acc)
print(null_auc)
print(lasso_auc)
print(rf_auc)
print(svm_auc)
```

```{r}
# Create a data frame with model names, accuracy, and AUC Based on test data
model_perfomance_test <- data.frame(
  Model = c("Null Model", "Lasso Model", "Random Forest Model", "Support Vector model"),  # Replace with your model names
  Accuracy = c(0.6023222	, 0.8026125	, 0.8156749		, 0.796807),  # Replace with your accuracy values
  AUC = c(0.5	, 0.8894864		, 0.8868349			, 0.882473			)  # Replace with your AUC values
)

# Print the results
print(model_perfomance_test)

# Save the table as a file

summarytable_file = here("results","tables", "model_perfomance_test.rds")
saveRDS(model_perfomance_test, file = summarytable_file)

```