### 2024-04-07 code/analysis-code/ordinary_logistic-analysis.qmd

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

## Introduction

This script uses tuberculosis(tb) as the outcome of interest and fits the following models to the analysis data:

* Null
* Decision Tree
* Random Forest
* LASSO

It compares the  models, and then finally fits the “best” model to the test data.

<br>

Follow the previous processing and analysis markdowns to generate the analysis data used here.

<br>

Each model will follow this process:

1. Model Specification
2. Workflow Definition
3. Tuning Grid Specification
4. Tuning Using Cross-Validation and the tune_grid() function
5. Identify Best Model
6. Model Evaluation

```{r seed=123}
defaultW <- getOption("warn") 
options(warn = -1) 
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(yardstick)
options(warn = defaultW)

```

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)
colnames(madaproject)
```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)
data_split <- initial_split(madaproject, prop = 0.7)
train <- training(data_split)
test <- testing(data_split)
ntrain = nrow(train)
ntest = nrow(test)
```

## ---- fit-data-logistic --------
# fit the logistic models with tb as outcome 
# second model has all variables as predictors
```{r}
log_mod <- logistic_reg() %>% set_engine("glm")
logmod2 <- log_mod %>% fit(tb ~ ., data = train)

```

```{r}
# Compute the accuracy and AUC for model 2
logmod2_acc <- logmod2 %>% 
  predict(train) %>% 
  bind_cols(train) %>% 
  metrics(truth = tb, estimate = .pred_class) %>% 
  filter(.metric %in% c("accuracy"))
```

```{r}
levels(train$tb)
```

```{r}

# Predict probabilities
logmod2_probs <- logmod2 %>%
  predict(new_data = train, type = "prob")  
```

```{r}
logmod2_auc <-  logmod2 %>%
  predict(new_data = train, type = "prob") %>%
  bind_cols(train) %>%
  roc_auc(truth = tb, .pred_YES)
```

```{r}
# Print the results
print(logmod2_acc)
print(logmod2_auc)
```

```{r}
# Bind predicted probabilities to the training data
logmod2_with_probs <- bind_cols(train, logmod2_probs)

#Save training data and predicted values as RDS
save_data_location <- here::here("data","processed-data","logmod2_with_probs.rds")
saveRDS(logmod2_with_probs, file = save_data_location)
```

```{r}
 #Calculate ROC curve
roc_curve_data <- yardstick::roc_curve(data = logmod2_with_probs, truth = tb, 
                                       .pred_YES)
#save the ROC curve as RDS

# Print ROC curve data
print(roc_curve_data)
```
```{r}
# Plot ROC curve
Roccurve <- ggplot(roc_curve_data, aes(x = sensitivity , y = 1 - specificity)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "blue") +
  labs(x = "False Positive Rate (TB CASES)", 
       y = "True Positive Rate (TB CASES)", 
       title = "ROC Curve") +
  theme_minimal()
Roccurve

figure_file = here("results","figures", "Roccurve.png")
ggsave(filename = figure_file, plot=Roccurve) 
```

#Model improvement

```{r}
#Set a seed
rngseed = 5002
set.seed(rngseed)

#Define cross-validation with k = 10 folds
folds <- vfold_cv(train, v = 10)

#setting model specification 
model_specf <- logistic_reg() %>% set_engine("glm")

# Workflow for all predictors
wfall4 <- workflow() %>%
  add_model(model_specf)%>%
  add_formula(tb ~ .,)

#Fitting the models with resampling 
fit2 <- fit_resamples(wfall4, resamples = folds)

#Getting metrics for both models 
collect_metrics(fit2)
```

`