### 2024-05-05 code/analysis-code/Cross_validation_Lasso.qmd
part61

---
title: "Predicting of the pulmunary TB cases , Uganda"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---

```{r}
# load the relevant tidymodels libraries
library(here) #for data loading/saving
library(tidyverse) #for data management
library(tidymodels) #for data modeling
library(skimr) #for variable summaries
library(broom.mixed) #for converting bayesian models to tidy tibbles
library(glmnet) #for lasso models
library(doParallel) #for parallel backend for tuning processes
library(ranger) #for random forest models
library(rsample)
library(vcd)# Load the vcd package for Cramer's V calculation
```

The purpose of this is to check if some variable are dropped by the lasso model so to get an insight of what variables

## Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","madaproject.rds")
#load data
# Loading the needed dataset
madaproject <- readRDS(data_location)

```

Before I start doing some model building I split the data into training and testing datasets.

```{r}
colnames(madaproject)
```

Now I do LASSO -Exploration

```{r}
set.seed(12345)
logistic_model <- logistic_reg() %>% set_engine("glm")

# Make sure you set reference level (the outcome you are NOT interested in)
levels(madaproject$tb)

data_cv10 <- vfold_cv(madaproject, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(tb ~ ., data = madaproject) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_model(logistic_lasso_spec_tune) %>%
    add_recipe(logistic_rec)

# Tune Model (trying a variety of values of Lambda penalty)
lasso_grid <- grid_regular(
  penalty(range = c(-3, 1)), 
  levels = 30)

tune_outputlasso <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE),
  grid = lasso_grid # lasso penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_outputlasso) + theme_classic()

# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_outputlasso, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the lowest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_lasso <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
fit(data = madaproject)
```

```{r}
# Get the best parameters
best_params <- show_best(tune_outputlasso, metric = "roc_auc") %>%
               slice(1)

# Finalize workflow with best parameters
final_lasso_wf <- finalize_workflow(log_lasso_wf, best_params)

# Fit model to training data
lasso_train_fit <- fit(final_lasso_wf, data = madaproject)

# Extract fitted model
fitted_model <- extract_fit_parsnip(lasso_train_fit)

# Extract selected variables
selected_variables <- tidy(fitted_model$fit) %>%
                      filter(!is.na(estimate) & estimate != 0) %>%
                      pull(term)

# Print selected variables
print(selected_variables)

```
I see that only one variable is dropped remaining with a total of 13 covariates.