### 2024-05-05 products/supplement/supplement.qmd

---
title: "CASE DETECTION OF TUBERCULOSIS PATIENTS USING MACHINE-LEARNING"
subtitle: ""
author: Patrick Kaggwa
date: "2024-02-24"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/MADAPROJECT.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(magrittr)
library(flextable)
library(htmltools)
```

## Supplemental Information
This document represents additional information about the data and results.

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic2
#| fig-cap: "TB_barplot"
#| echo: false
knitr::include_graphics(here("results","figures", "tb_barplot.png"))
```
```{r echo=FALSE, message = FALSE}
#| label: fig-schematic3
#| fig-cap: "Age distribution"
#| echo: false
knitr::include_graphics(here("results","figures", "age_distribution.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic10
#| fig-cap: "AGe _TB Box PLot"
#| echo: false
knitr::include_graphics(here("results","figures", "age_boxplot.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: fig-schematic4
#| fig-cap: "Hivstatus barplot"
#| echo: false
knitr::include_graphics(here("results","figures", "hivstatus_barplot.png"))
```

```{r echo=FALSE, message = FALSE}
#| label: tbl-schematic8
#| tbl-cap: "Table showing Characteristics"
ls<- readRDS(here("results", "tables", "characteristics.rds"))
ls %>%
  knitr::kable()
```